Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability

### 2025-10-16 paper/techdebt2026/techdebt.tex
version with 5 + 1 pages and new title
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\title{Mitigating Code Duplication in the Linux Kernel}

@@

\begin{figure}[ht]
\centering
\includegraphics[scale=0.325]{timeline_patch}
\caption{Timeline of the PATCH submission process}
\label{fig:timeline}
\end{figure}

The experience of submitting a patch was not as we expected. The process took 7 months and 16 days, significantly slower than expected. We initially submitted the patch on August 9, 2024, but we did not receive replies. Thus, we had to resend it on October 9, 2024. We received an initial response on November 3, 2024, requesting minimal changes. We submitted a second version with the requested changes on November 16, 2024. We did not receive replies initially and will need to resend the patch on February 7, 2025. We received a reply on February 11, 2025, asking for additional changes. We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025. Figure \ref{fig:timeline} illustrates the timeline of the patch submission.

After the initial reply from the maintainers, we observed that the review process became more responsive. We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers. Another contributing factor to the initial delay could be the maintainers' workload at the time of the first submission.

We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.

Regarding the patch changes requested for the second version, the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown. For the change requested on the third version, the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.

@@

\begin{table*}
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ | c | m{15mm} | m{20mm} | m{30mm} | m{30mm} | m{20mm} | }
\hline
\textbf{Group} & \textbf{Subsystem} & \textbf{Similarity} & \textbf{Patch Status} & \textbf{Refactoring Method} & \textbf{Code Diff} \\
\hline

1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & Parameterize method & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & Parameterize method & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & Parameterize method & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & Parameterize method & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & Extract method & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & Extract method & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) -- Partially & Parameterize method & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & Inline method & +2/-12 \\ 
   & AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline

@@

\label{tab:stu}
\end{table*}

@@

The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by the \texttt{anonymous-tool} tool. The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}.

>>>>

\title{Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability}

@@

%\begin{figure}[ht]
%\centering
%\includegraphics[scale=0.325]{timeline_patch}
%\caption{Timeline of the PATCH submission process}
%\label{fig:timeline}
%\end{figure}

%The experience of submitting a patch was not as we expected. The process took 7 months and 16 days, significantly slower than expected. We initially submitted the patch on August 9, 2024, but we did not receive replies. Thus, we had to resend it on October 9, 2024. We received an initial response on November 3, 2024, requesting minimal changes. We submitted a second version with the requested changes on November 16, 2024. We did not receive replies initially and will need to resend the patch on February 7, 2025. We received a reply on February 11, 2025, asking for additional changes. We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
%Figure \ref{fig:timeline} illustrates the timeline of the patch submission.

%After the initial reply from the maintainers, we observed that the review process became more responsive. We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers. Another contributing factor to the initial delay could be the maintainers' workload at the time of the first submission.

%We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that

We submitted the initial patch: the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown. We submitted a second version: the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one. We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.

During this contribution process, we learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.

@@

\begin{table}
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ |c |c |c |c |c | c| }
\hline
\textbf{G} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Method} & \textbf{Diff} \\
\hline

1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & Parameterize & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & Parameterize & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & Parameterize & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & Parameterize & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & Extract & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & Extract & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted* (v6) & Parameterize & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & Inline & +2/-12 \\ 
   & AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline

@@

\label{tab:stu}
\end{table}

@@

The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by the \texttt{anonymous-tool} tool. The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches.
%TODO: disponibilizar artefato da mesma forma que mandamos para a ICSE Education Track~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}
